Automatic Identification and Annotation of Tasks in Structured Programs

### 2018-01-26 docs/ACM/paper.tex
added section 3.1

The algorithm to find tasks is illustrated in Figure \ref{fig:alg_main}. It uses
information of two abstract program structures: the \emph{Control Flow Graph (CFG)} and
the \emph{Program Dependence Graph (PDG)}. The key intuition behind
finding potential tasks relies on a concept that we call \emph{Helices}.

\begin{definition}[Helix]
\label{def:helix}
Given a set $S$ of strongly connected nodes in the Dependence Graph $G$, a set of nodes $H$ is a Helix if, and only if:
\begin{itemize}
\item $H \subset \bigcup_i N_i[S]$;
\item $H  \cap S = \emptyset$;
\item $H$ is connected and maximal.
\end{itemize}
\end{definition}

We can view a \emph{Helix} as a set of nodes in the PDG that depends on a Strongly Connected Component
(SCC) but \emph{do not integrate the SCC}. That means helices are all those subgraphs that depend on
the SCC and do not form a cyclic dependence with the SCC itself.

Statically, a Strongly Connected Component in the PDG represents iteration dependence, so it
arises due to loops or recursive patterns, should the PDG is interprocedural. The insight behind 
\emph{Helices} is that they represent instructions that are deemed to be executed more than once
for they depend on a cyclic dependence (the SCC), and, since they do not integrate the SCC, they can
be executed in parallel.

Thus, the first step in the algorithm to find tasks is to map these \emph{Helices}
into program regions. The algorithm goes through the Dependence Graph looking for \emph{Helices}
and then finds the minimal covering region for each \emph{Helix}. These minimal regions
are later going to be annotated as tasks in the original program, should they fit a cost model.

@@

\begin{figure}[t!]
\begin{center}
\includegraphics[width=1\columnwidth]{images/expand_alg}
\caption{Task discovery via expansion of hammock regions.}
\label{fig:expand_alg}
\end{center}
\end{figure}

@@

@@

@@

@@

>>>>

The {\Taskminer} algorithm is illustrated in FIgure \ref{fig:alg_main}. 
It uses information from both the \emph{Control Flow Graph} and
the \emph{Dependence Graph}. It relies on the concepts of
\emph{windmills} and \emph{helices} (Definitions \ref{def:helix} and \ref{def:windmill}).

\begin{definition}[Windmill]
\label{def:windmill}
Given a set $S$ of strongly connected nodes in the program \emph{Dependence Graph} $G$, 
a windmill $W$ is defined as:
$$W = \bigcup_i N_i[S]$$
\end{definition}

Statically, a strongly connected set $S$ in the  \emph{Dependence Graph} $G$ represents cyclic dependence between instructions in any iterative construct.
It arises from a loop in the program or, if $G$ is interprocedural,
from a recursive pattern.

Let $L = \{S_{1}, S_{2}, ..., S_{n}\} \subset W$ be the list of $n$ strongly connected sets contained
in a windmill $W$. We call $S_{b}$ to be a \emph{windmill base} if $\forall S_{i} \subset L$, $S_{b} \geq S_{i}$ in any topological sorting of $G$.

\begin{definition}[Helix]
\label{def:helix}
Given a windmill $W$  and a windmill base $S \subset W$ in the \emph{Dependence Graph} $G$, a set of nodes $H$ is a Helix if, and only if:
\begin{itemize}
\item $H \subset W$;
\item $H  \cap S_{b} = \emptyset$;
\item $H$ is connected and maximal.
\end{itemize}
\end{definition}

We can view a \emph{Helix} as a set of nodes in the \emph{Dependence Graph}
that depends on a strongly connected set which is a \emph{windmill base}.
That means helices are all those subgraphs that depend on
the \emph{windmill base} but do not form a cyclic dependence with it.
Figure \ref{fig:helix} illustrates both concepts.

The insight behind  \emph{Helices} is that they represent instructions that 
are deemed to be executed more than once, for they depend on a cyclic
dependence (\emph{windmill base}), and, since they do not form a strongly
connected set itself, they can be executed in parallel.

\begin{figure}[t!]
\begin{center}
\includegraphics[width=1\columnwidth]{images/helix}
\caption{\emph{Windmills} and \emph{Helices} corresponding to the sample code above.
We see the \emph{windmill} 
$W = \{A, B, C, D, E, F, G, H\}$, the windmill base $S_{b} = \{A, B, C\}$ 
and two \emph{helices} $H_{1} = \{H\}$ and $H_{2} = \{D, E, F, G\}$.  Note
that there are more than one strongly connected, but only of them is a
base, which is the one at the top topologically.}
\label{fig:helix}
\end{center}
\end{figure}

Thus, the first step in the algorithm to find tasks is to map these \emph{Helices}
into program regions. The algorithm scans through the \emph{Dependence Graph}
looking for \emph{Helices} and then finds the minimal covering region for each \emph{Helix}.
These minimal regions are later going to be annotated as tasks in the original
program, should they fit a cost model.

@@

Figure \ref{fig:expand_alg} illustrate the algorithm of expansion. It simply tries to expand while
all the conditions in Definition \ref{def:expansion} are met, and then halts
when they are not. After this process, {\Taskminer} ends up with a list
of potential \emph{task regions}, which will then pass under
a series of operations prior to the annotation to source code, the 
ultimate step.

\begin{figure}[h]
\begin{center}
\includegraphics[width=1\columnwidth]{images/expand_alg}
\caption{Task discovery via expansion of hammock regions.}
\label{fig:expand_alg}
\end{center}
\end{figure}

@@

%I don't know much of this part. I asked gleison to write me a paragraph and
%then I'll review it and try to write something here.

@@

%We need to discuss the THRESHOLD. Are we going to use
%César's experiments at Unicamp concerning the different runtimes? 
%Because if yes we need to think on how to write it.

@@

%I can work on that on the weekend!

@@

%I asked gleison and he'll sit with me to explain everything ,
%how it's done, and then I'll write it here.
